---
title: "Chapter 6 Chaz Clark"
output:
  html_document:
    df_print: paged
---
```{r}
library(fpp2)
plastics
```

6.2The plastics data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.
a.Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle?
```{r}
autoplot(plastics)
```
ANSWER:THere is well defined cycles in the data, as well as a generally linear up trend.

b.Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.
```{r}
plastics %>% decompose(type = "multiplicative") %>% 
  autoplot() + xlab("Year") +
  ggtitle("Classical multiplicative decomposition of plastics")
```

c.Do the results support the graphical interpretation from part a?
ANSWER: yes the initial assessment from a matches the charts produced in b

d.Compute and plot the seasonally adjusted data.
```{r}
library(seasonal)
library(xts)
plastics <- ts(plastics,frequency = 12, start = 2001)
#note the years of the data are not in the description but seas() has a >year999 error this was my way to get arround it
plastics %>% seas() -> fit
fit
```
```{r}
autoplot(plastics, series="Data") +
  autolayer(trendcycle(fit), series="Trend") +
  autolayer(seasadj(fit), series="Seasonally Adjusted") +
  xlab("Year") + ylab("New orders index") +
  ggtitle("Plastics Manufacturing") +
  scale_colour_manual(values=c("gray","blue","red"),
             breaks=c("Data","Seasonally Adjusted","Trend"))
```

e.Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?
```{r}
plastics[12]
plastics[12]<-plastics[12]+500
plastics[12]
```
```{r}
plastics %>% seas() -> fit
fit
```
```{r}
autoplot(plastics, series="Data") +
  autolayer(trendcycle(fit), series="Trend") +
  autolayer(seasadj(fit), series="Seasonally Adjusted") +
  xlab("Year") + ylab("New orders index") +
  ggtitle("Plastics Manufacturing") +
  scale_colour_manual(values=c("gray","blue","red"),
             breaks=c("Data","Seasonally Adjusted","Trend"))
```
ANSWER: the outlier had a large affect on the seasnally adjusted data model.

f.Does it make any difference if the outlier is near the end rather than in the middle of the time series?
```{r}
plastics[30]<-plastics[30]+500
plastics[50]<-plastics[50]+500
plastics %>% seas() -> fit
autoplot(plastics, series="Data") +
  autolayer(trendcycle(fit), series="Trend") +
  autolayer(seasadj(fit), series="Seasonally Adjusted") +
  xlab("Year") + ylab("New orders index") +
  ggtitle("Plastics Manufacturing") +
  scale_colour_manual(values=c("gray","blue","red"),
             breaks=c("Data","Seasonally Adjusted","Trend"))
```
ANSWER: It appear that the outlier near the eand has a higher impact due to the trend having a higher trend at tehat time.

6.4 Figures 6.16 and 6.17 show the result of decomposing the number of persons in the civilian labour force in Australia each month from February 1978 to August 1995.

Seasonal component from the decomposition shown in the previous figure.
Figure 6.17: Seasonal component from the decomposition shown in the previous figure.

a. Write about 3–5 sentences describing the results of the decomposition. Pay particular attention to the scales of the graphs in making your interpretation.
ANSWER:
1. the data is very linear in its trend upward, with some variation
2. seasonality seems strong and increasing over time
3. the volitility of the seasonality seems to increase over time, higher highs and lower lows

Note: i am not sure how to read the scales for the seaspnal sub-series plot to make sense of it just yet

b. Is the recession of 1991/1992 visible in the estimated components?
ANSWER: yes there is a dip in the data and the remainder is significantly strong in 1991. If a dummy variable was created for these months it would be interesting to see if that explains that whold section of data or not.