---
title: "Chapter 7.5 & 7.6 Chaz Clark"
output:
  html_document:
    df_print: paged
---
```{r}
library(fpp2)
head(books)
```

5. Data set books contains the daily sales of paperback and hardcover books at the same store. The task is to forecast the next four days’ sales for paperback and hardcover books.
a.Plot the series and discuss the main features of the data.
```{r}
autoplot(books)
#The values of both series have and upward trend within the same band
#The Paperback appears to have some sort of trend
#The Hard Cover pattern is less trend looking then paperback
#both trends look linear so no log() is needed on the data
```

b.Use the ses() function to forecast each series, and plot the forecasts.
```{r}
#this is where i was going down the wrong path...
#library(seasonal)
#library(lubridate)
#books_fc <- ts(books, frequency = 7, start = decimal_date(as.Date("2020-01-01")))
#books_fc[,1] %>% decompose(type = "multiplicative") %>% 
#  autoplot() 
#seas() cannot work for daily data, frequency must be monthly or quarterly
#frequency =12,30,365,7 gives error
#books_fc[,1] %>% seas(x11="") -> fit
#fit %>% autoplot()

#thanks to your help:
fcast1 <- ses(books[,"Hardcover"],h=4)
fcast2 <- ses(books[,"Paperback"],h=4)
autoplot(books) + 
       autolayer(fcast1, series="Hardcover",PI=FALSE) + 
       autolayer(fcast2, series="Paperback",PI=FALSE)
```

c.Compute the RMSE values for the training data in each case.
```{r}
round(accuracy(fcast1),2)
round(accuracy(fcast2),2)
```

6.We will continue with the daily sales of paperback and hardcover books in data set books.
a.Apply Holt’s linear method to the paperback and hardback series and compute four-day forecasts in each case.
```{r}
hfcast1 <- holt(books[,"Hardcover"],h=4)
hfcast2 <- holt(books[,"Paperback"],h=4)
autoplot(books) + 
       autolayer(hfcast1, series="Hardcover",PI=FALSE) + 
       autolayer(hfcast2, series="Paperback",PI=FALSE)
```

b.Compare the RMSE measures of Holt’s method for the two series to those of simple exponential smoothing in the previous question. (Remember that Holt’s method is using one more parameter than SES.) Discuss the merits of the two forecasting methods for these data sets.
```{r}
#hardcover ses()
round(accuracy(fcast1),2)
#paperback ses()
round(accuracy(fcast2),2)
#hardcover holt()
round(accuracy(hfcast1),2)
#paperback holt()
round(accuracy(hfcast2),2)
#The RMSE for ses() are both higher than either for holt()
#The extra 'trend' that holt uses seems to lower its RSME
#Honestly at this point i am not sure of the merits of the two methods for these data sets.
```

c.Compare the forecasts for the two series using both methods. Which do you think is best?
```{r}
#Visually i like the results of holt better. The trend carries a lot of weight with it.
#I actually think the ses() is better for Hardcover and the Holt better for paperback.
#That said the numbers are all close together and puts them all in the same ball park. 
#However strict comparison of the RMSE puts SES as it has higher scores then ses for both forecasts.
#Other than that i dont have a great grasp of these values yet and how to compare them, knowing some you cant.
```

d.Calculate a 95% prediction interval for the first forecast for each series, using the RMSE values and assuming normal errors. Compare your intervals with those produced using ses and holt.
```{r}
fit1 <- ets(books[,"Hardcover"])
#Holt RMSE 27.19
#SES  RMSE 31.93
#ETS  RSME 27.22
fit2 <- ets(books[,"Paperback"])
#Holt RMSE 21.14
#SES  RMSE 33.64
#ETS  RSME 33.64
summary(fit1)
summary(fit2)
#it appears that SES is a better forecast for Hardcovers
# and ETS and SES tie for Paperback. This may mean that 
#ETS used SES for its results.(ETS(A,N,N))
```