---
title: "Chapter 8.8-8.9 Chaz Clark"
output: html_notebook
---
8.Consider austa, the total international visitors to Australia (in millions) for the period 1980-2015.
```{r}
#load the libraries and data
library(fpp2)
austa
```
a. Use auto.arima() to find an appropriate ARIMA model. What model was selected. Check that the residuals look like white noise. Plot forecasts for the next 10 periods.
```{r}
austa %>% auto.arima() %>% checkresiduals()
#Ljung-Box test
#
#data:  Residuals from ARIMA(0,1,1) with drift
#Q* = 2.297, df = 5, p-value = 0.8067
#
#Model df: 2.   Total lags used: 7

#residuals look like white noise, the ACF also looks well within bounds
```
```{r}
austa %>% auto.arima() %>% forecast(h=10) %>% autoplot()
```
b. Plot forecasts from an ARIMA(0,1,1) model with no drift and compare these to part a. Remove the MA term and plot again.
```{r}
austa %>% Arima(order = c(0,1,1)) %>% forecast(h=10) %>% autoplot()
```
```{r}
austa %>% Arima(order = c(0,1,0)) %>% forecast(h=10) %>% autoplot()
```
c. Plot forecasts from an ARIMA(2,1,3) model with drift. Remove the constant and see what happens.
```{r}
austa %>% Arima(order = c(2,1,3),include.drift = 5) %>% forecast(h=10) %>% autoplot()
```
```{r}
austa %>% Arima(order = c(2,1,3)) %>% forecast(h=10) %>% autoplot()
```
d. Plot forecasts from an ARIMA(0,0,1) model with a constant. Remove the MA term and plot again.
```{r}
austa %>% Arima(order = c(0,0,1),include.drift = 2) %>% forecast(h=10) %>% autoplot()
```
e. Plot forecasts from an ARIMA(0,2,1) model with no constant.
```{r}
austa %>% Arima(order = c(0,0,0),include.drift = 2) %>% forecast(h=10) %>% autoplot()
```

9.For the usgdp series:
```{r}
usgdp %>% autoplot()
```
a. if necessary, find a suitable Box-Cox transformation for the data;
```{r}
#data is not linear and curves indicating a log or Box-Cox is needed
#usgdp %>% log() %>%  nsdiffs()
#[1] 0
#indicates that no seasonality is in the data
#usgdp %>% log() %>% ndiffs()
#[1] 1
#indicates 1 diff of the data is needed

#usgdp %>% diff() %>% Box.test(type="Ljung-Box")
#Box-Ljung test
#
#data:  .
#X-squared = 39.187, df = 1, p-value = 3.85e-10
#this indicates the model has sufficient differences

cbind("Original Data" = usgdp,
      "BoxCox" = BoxCox(usgdp, lambda = 0),
      #"Seasonally\n differenced logs" =
        #diff(log(usgdp),12),
      "Single\n differenced logs" =
        diff(BoxCox(usgdp, lambda = 0),1)) %>%
  autoplot(facets=TRUE) +
    xlab("Year") + ylab("") 
#the chart below shows the BoxCox straightened the data, and the single diff() was sufficient to produce white noise 
```
b. fit a suitable ARIMA model to the transformed data using auto.arima();
```{r}
BoxCox(usgdp, lambda = 0) %>% auto.arima() 
```
c. try some other plausible models by experimenting with the orders chosen;
```{r}
BoxCox(usgdp, lambda = 0) %>% Arima(order = c(0,1,1))
#ARIMA  AIC       AICboxcox 
#212    2437.32   -1497.48
#010    2619.01   -1379.59
#110    2476.09   -1488.12
#011    2549.51   -1441.45
#AIC indicates 212 has the lowest value
#AICboxcox indicates 212 as well
```
d. choose what you think is the best model and check the residual diagnostics;
```{r}
BoxCox(usgdp, lambda = 0) %>% Arima(order = c(2,1,2)) %>% checkresiduals()
#residuals look good, appears to be white noise, and the ACF is mostly within range
```
e. produce forecasts of your fitted model. Do the forecasts look reasonable?
```{r}
usgdp %>% Arima(order = c(2,1,2), include.drift = 0.0084, lambda = 0) %>% forecast(h=20) %>% autoplot()
#the forecast does seem reasonable but the data is still in log form
#BoxCox(usgdp, lamda = 0) %>% Arima(order = c(2,1,2), include.drift = 0.0084) %>% forecast(h=20) %>% autoplot()
#Q - should i revert it back here prior to forecasting it?
#A - if you fix the calculation so Arima has lambda then its a lot easier and it reverts the log for you
```
f. compare the results with what you would obtain using ets() (with no transformation).
```{r}
usgdp %>%  ets() %>%  forecast(h=20) %>% autoplot()
```
```{r}
#usgdp %>% ets() %>% forecast(h=20) %>%  accuracy() 
#                  ME     RMSE      MAE        MPE      MAPE     MASE       ACF1
#Training set 1.30185 41.53448 31.85324 0.02319607 0.7543931 0.180026 0.07596492

usgdp %>% Arima(order = c(2,1,2), include.drift = 0.0084, lambda = 0) %>% accuracy()
#                    ME     RMSE      MAE         MPE      MAPE      MASE        ACF1
#Training set -1.181516 39.45022 29.23715 0.001992878 0.6710562 0.1652405 -0.01317471

#the ARIMA RMSE is slightly better than the ETS, it would be worth checking the residuals for both and comparing with only a 2 point difference
```