---
title: "Chapter 8.1-8.3, 8.5 Chaz Clark"
output: html_notebook
---
Figure 8.31 shows the ACFs for 36 random numbers, 360 random numbers and 1,000 random numbers.
a. Explain the differences among these figures. Do they all indicate that the data are white noise?
```{r}
#Yes there are no significant Lags outside the bounds of the blue lines so statistically not significant, or no correlation in the data between lags
```
b.Why are the critical values at different distances from the mean of zero? Why are the autocorrelations different in each figure when they each refer to white noise?
```{r}
#The critical values differ based on what is statistically significant at a 95% level, where these three have differeing size of datasets they will have different values.
#White noise signifies a static dataset, but not all white noise is created equal. The sequence of the random generated nubmers will play a factor in the lags of the data, and thus each will ahve a different looking white noise.
```

2.A classic example of a non-stationary series is the daily closing IBM stock price series (data set ibmclose). Use R to plot the daily closing prices for IBM stock and the ACF and PACF. Explain how each plot shows that the series is non-stationary and should be differenced.
```{r}
library(fpp2)
#ibmclose
autoplot(ibmclose)
#you can see small repeating patterns in the chart indicating is non-stationary
```
```{r}
ggAcf(ibmclose)
#here you can see that in the ACF plot all lags are outside the blue lines indicating non-stationary
```
```{r}
ggPacf(ibmclose)
#in the PACF plot Lag1 is significantly outside of the lines, this too indicates non-stationary
```

3.For the following series, find an appropriate Box-Cox transformation and order of differencing in order to obtain stationary data.
a. usnetelec
```{r}
#usnetelec  %>% nsdiffs()
#Error in nsdiffs(.) : Non seasonal data
#usnetelec  %>% log() %>% ndiffs()
#[1] 2
#this indicates no seasonality but that two levels of diff() is required for stationary data

#usnetelec %>% diff() %>% Box.test(type="Ljung-Box")
#X-squared = 0.8508, df = 1, p-value = 0.3563
#This contradicts ndiffs() indicates that diff() is needed once 

#library(urca)
#usnetelec %>% ur.kpss() %>% summary()
#usnetelec %>% diff() %>%  ur.kpss() %>% summary()
#this is less than 1% so 1 diff is enough

####################### 
# KPSS Unit Root Test # 
####################### 

#Test is of type: mu with 3 lags. 
#
#Value of test-statistic is: 0.1585 
#
#Critical value for a significance level of: 
#                10pct  5pct 2.5pct  1pct
#critical values 0.347 0.463  0.574 0.739

#the chart below confirms white noise with a single diff()
cbind("Original Data" = usnetelec,
      "BoxCox" = BoxCox(usnetelec, lambda = 0),
      #"Seasonally\n differenced logs" =
        #diff(log(usnetelec),12),
      "Single\n differenced logs" =
        diff(BoxCox(usnetelec, lambda = 0),1)) %>%
  autoplot(facets=TRUE) +
    xlab("Year") + ylab("") 
```
b. usgdp
```{r}
#usgdp  %>% nsdiffs()
#[1] 0
#usgdp  %>% log() %>% ndiffs()
#[1] 1
#this indicates no seasonality but that one levels of diff() is required for stationary data

#usgdp %>% diff() %>% Box.test(type="Ljung-Box")
#X-squared = 39.187, df = 1, p-value = 3.85e-10
#This also indicates that diff() is needed once 

#library(urca)
#usgdp %>% ur.kpss() %>% summary()
#sgdp %>% diff() %>%  ur.kpss() %>% summary()
#this is higher than 1% but under 2, in this case i defer to the previous tests

####################### 
# KPSS Unit Root Test # 
####################### 

#Test is of type: mu with 4 lags. 
#
#Value of test-statistic is: 1.7909 
#
#Critical value for a significance level of: 
#                10pct  5pct 2.5pct  1pct
#critical values 0.347 0.463  0.574 0.739

#the chart below signals white noise after 1 diff()
cbind("Original Data" = usgdp,
      "BoxCox" = BoxCox(usgdp, lambda = 0),
      #"Seasonally\n differenced logs" =
        #diff(log(usgdp),12),
      "Single\n differenced logs" =
        diff(BoxCox(usgdp, lambda = 0),1)) %>%
  autoplot(facets=TRUE) +
    xlab("Year") + ylab("") 
```
c. mcopper
```{r}
#mcopper  %>% nsdiffs()
#[1] 0
#mcopper  %>% log() %>% ndiffs()
#[1] 1
#this indicates no seasonality but that one levels of diff() is required for stationary data

#mcopper %>% diff() %>% Box.test(type="Ljung-Box")
#X-squared = 38.911, df = 1, p-value = 4.435e-10
#This also indicates that diff() is needed once 

#library(urca)
#mcopper %>% ur.kpss() %>% summary()
#mcopper %>% diff() %>%  ur.kpss() %>% summary()
#this is less than 1% so only 1 diff() is needed

####################### 
# KPSS Unit Root Test # 
####################### 

#Test is of type: mu with 6 lags. 
#
#Value of test-statistic is: 0.1843 
#
#Critical value for a significance level of: 
#                10pct  5pct 2.5pct  1pct
#critical values 0.347 0.463  0.574 0.739

#the chart below signals white noise after 1 diff() 
cbind("Original Data" = mcopper,
      "BoxCox" = BoxCox(mcopper, lambda = 0),
      #"Seasonally\n differenced logs" =
        #diff(log(mcopper),12),
      "Single\n differenced logs" =
        diff(BoxCox(mcopper, lambda = 0),1),
      "Double\n difference  logs" = 
        diff(diff(log(mcopper),12),1)) %>%
  autoplot(facets=TRUE) +
    xlab("Year") + ylab("") 
```
d. enplanements
```{r}
#enplanements  %>% nsdiffs()
#[1] 1
#enplanements  %>% log() %>% ndiffs()
#[1] 1
#this indicates one diff() for seasonality and that one levels of diff() is required for stationary data

#enplanements %>% diff() %>% Box.test(type="Ljung-Box")
#X-squared = 12.948, df = 1, p-value = 0.0003203
#This also indicates that diff() is needed once 

#library(urca)
#enplanements %>% ur.kpss() %>% summary()
#enplanements %>% diff() %>%  ur.kpss() %>% summary()
#this is less than 1% so only 1 diff() is needed

####################### 
# KPSS Unit Root Test # 
####################### 

#Test is of type: mu with 5 lags. 
#
#Value of test-statistic is: 0.0086 
#
#Critical value for a significance level of: 
#                10pct  5pct 2.5pct  1pct
#critical values 0.347 0.463  0.574 0.739

#the chart below signals white noise after 1 diff() seasonal and 1 diff() non-seasonal
cbind("Original Data" = enplanements,
      "BoxCox" = BoxCox(enplanements, lambda = 0),
      "Seasonally\n differenced logs" =
        diff(log(enplanements),12),
      "Single\n differenced logs" =
        diff(BoxCox(enplanements, lambda = 0),1),
      "Double\n difference  logs" = 
        diff(diff(log(enplanements),12),1)) %>%
  autoplot(facets=TRUE) +
    xlab("Year") + ylab("") 
```
e. visitors
```{r}
#visitors  %>% nsdiffs()
#[1] 1
#visitors  %>% log() %>% ndiffs()
#[1] 1
#this indicates no seasonality but that one levels of diff() is required for stationary data

#visitors %>% diff() %>% Box.test(type="Ljung-Box")
#X-squared = 29.101, df = 1, p-value = 6.871e-08
#This also indicates that diff() is needed once 

#library(urca)
#visitors %>% ur.kpss() %>% summary()
#visitors %>% diff() %>%  ur.kpss() %>% summary()
#this is less than 1% so only 1 diff() is needed

####################### 
# KPSS Unit Root Test # 
####################### 

#Test is of type: mu with 4 lags. 
#
#Value of test-statistic is: 0.0191 
#
#Critical value for a significance level of: 
#                10pct  5pct 2.5pct  1pct
#critical values 0.347 0.463  0.574 0.739

#the chart below signals white noise after 1 diff() seasonal and 1 diff() non-seasonal
cbind("Original Data" = visitors,
      "BoxCox" = BoxCox(visitors, lambda = 0),
      "Seasonally\n differenced logs" =
        diff(log(visitors),12),
      "Single\n differenced logs" =
        diff(BoxCox(visitors, lambda = 0),1),
      "Double\n difference  logs" = 
        diff(diff(log(visitors),12),1)) %>%
  autoplot(facets=TRUE) +
    xlab("Year") + ylab("") 
```

5.For your retail data (from Exercise 3 in Section 2.10), find the appropriate order of differencing (after transformation if necessary) to obtain stationary data.
```{r}
library(readxl)
retail <- readxl::read_excel("~/Documents/Documents - Mac Pro/Elmhurst/MDS 570 Time Series/Week4/retail.xlsx", 
    skip = 1)
myts <- ts(retail[,"A3349873A"],
  frequency=12, start=c(1982,4))

#myts  %>% nsdiffs()
#[1] 1
#myts  %>% log() %>% ndiffs()
#[1] 1
#this indicates no seasonality but that one levels of diff() is required for stationary data

#myts %>% diff() %>% Box.test(type="Ljung-Box")
#X-squared = 34.684, df = 1, p-value = 3.878e-09

#This also indicates that diff() is needed once 

#library(urca)
#myts %>% ur.kpss() %>% summary()
#myts %>% diff() %>%  ur.kpss() %>% summary()
#this is less than 1% so only 1 diff() is needed

####################### 
# KPSS Unit Root Test # 
####################### 

#Test is of type: mu with 5 lags. 
#
#Value of test-statistic is: 0.0522 
#
#Critical value for a significance level of: 
#                10pct  5pct 2.5pct  1pct
#critical values 0.347 0.463  0.574 0.739

#the chart below signals white noise after 1 diff() seasonal and 1 diff() non-seasonal
cbind("Original Data" = myts,
      "BoxCox" = BoxCox(myts, lambda = 0),
      "Seasonally\n differenced logs" =
        diff(log(myts),12),
      "Single\n differenced logs" =
        diff(BoxCox(myts, lambda = 0),1),
      "Double\n difference  logs" = 
        diff(diff(log(myts),12),1)) %>%
  autoplot(facets=TRUE) +
    xlab("Year") + ylab("") 
```