---
title: "Stock Data Forecasting Chaz Clark"
output: html_notebook
---

Choose some time series data of interest to you, perhaps data that might be applicable to your course project. Use some of the exponential smoothing methods you have been practicing to produce an initial forecast for these data. Use this forum to discuss your solutions, questions, and ideas.
```{r}
#load the libraries
library(quantmod)
library(xts)
library(data.table)
library(tidyr)
library(dplyr)
library(gapminder)
library(TTR)
library(plyr)
library(fpp2)
```

```{r}
#get the data
options("getSymbols.warning4.0"=FALSE)
getSymbols('AAPL', from = "2007-01-03", to = "2020-5-14")
autoplot(AAPL$AAPL.Close)
#tail(AAPL)
AAPL2<-AAPL

getSymbols('AAPL', from = "2007-01-03", to = "2020-5-13")
AAPL3<-AAPL

getSymbols('AAPL', from = "2007-01-03", to = "2020-5-8")
AAPL4<-AAPL

getSymbols('AAPL', from = "2007-01-03", to = "2020-4-30")
AAPL5<-AAPL

#convert to dt
AAPLdt <- data.table(Date=index(AAPL2), coredata(AAPL2))

#convert to weekly
AAPLw <- to.weekly(AAPLdt)
#convert to monthly
AAPLm <- to.monthly(AAPLdt)

#convert to dt
AAPLdt2 <- data.table(Date=index(AAPL3), coredata(AAPL3))
AAPLdt3 <- data.table(Date=index(AAPL4), coredata(AAPL4))
AAPLdt4 <- data.table(Date=index(AAPL5), coredata(AAPL5))
#convert to weekly
AAPLw2 <- to.weekly(AAPLdt3)
#convert to monthly
AAPLm2 <- to.monthly(AAPLdt4)


#library(padr)
#AAPL2 <- AAPLdt %>% pad()

#set some variables
#i=4
df = 5
wf = 52
mf = 12

#library(timeDate)
# A timeDate Sequence
#date.sequence <- timeSequence(as.Date("2007-01-03"), as.Date("2020-02-28"));  # a short example period with three London holidays
#date.sequence;

# holidays in the period
#years.included <- unique( as.integer( format( x=date.sequence, format="%Y" ) ) );
#holidays <- holidayNYSE(years.included)  #  (locale was not specified by OP in question nor in profile, so this assumes for example: holidayLONDON; also supported by timeDate are: holidayNERC, holidayNYSE, holidayTSX & holidayZURICH)

# Subset business days
#business.days <- date.sequence[isBizday(date.sequence, holidays)]; 
#business.days

#source:https://stackoverflow.com/questions/13673895/r-time-series-data-daily-only-working-days

#convert dt to ts with xts
#mytsd2 <- ts(AAPLdt[,"AAPL.Close"])
#mytsd2
#mytsw2 <- xts(AAPLw[,"AAPLdt.Close"], order.by = as.POSIXct(business.days))
#mytsm2 <- xts(AAPLm[,"AAPLdt.Close"], order.by = as.POSIXct(business.days))

#convert dt to ts
mytsd <- ts(AAPLdt[,"AAPL.Close"], frequency=df, start=c(2007,1))
mytsw <- ts(AAPLw[,"AAPLdt.Close"], frequency=wf, start=c(2007,1))
mytsm <- ts(AAPLm[,"AAPLdt.Close"], frequency=mf, start=c(2007,1))

#set the training datasets
mytsd.train <- ts(AAPLdt2[,"AAPL.Close"], frequency=df, start=c(2007,1))
mytsw.train <- ts(AAPLw2[,"AAPLdt3.Close"], frequency=wf, start=c(2007,1))
mytsm.train <- ts(AAPLm2[,"AAPLdt4.Close"], frequency=mf, start=c(2007,1))

#forecast paramaters
hd = length(mytsd) - length(mytsd.train)
hw = length(mytsw) - length(mytsw.train)
hm = length(mytsm) - length(mytsm.train)
```

```{r}
#start with daily
#ETSd <- mytsd.train %>% ets(model="ZZZ", lambda = 0, biasadj = TRUE, allow.multiplicative.trend = TRUE) %>% forecast(h=hd)
ETSd <- mytsd.train %>% ets() %>% forecast(h=hd)
#ARIMAd1 <- mytsd.train %>% auto.arima(lambda = 0,biasadj = TRUE, stepwise = FALSE, approximation = FALSE) %>% forecast(h=hd)
ARIMAd <- mytsd.train %>% auto.arima(lambda = 0,biasadj = TRUE) %>% forecast(h=hd)
trend <- seq_along(mytsd.train)
#ARIMAd2 <- mytsd.train %>% auto.arima(lambda = 0,biasadj = TRUE, d=0, xreg=trend) %>% forecast(h=hd, xreg = length(mytsd.train))
STLd <- mytsd.train  %>%  stlf(lambda = 0, h=hd, biasadj = TRUE)
MSTLd <- mytsd.train  %>%  mstl(lambda = 0) %>% forecast(h=hd)
NNARd <- mytsd.train %>%  nnetar(lambda = 0) %>% forecast(h=hd)
TBATSd <- mytsd.train %>% tbats(biasadj = TRUE) %>% forecast(h=hd)
Combinationd <- (ETSd[["mean"]] + ARIMAd[["mean"]] + STLd[["mean"]] +  NNARd[["mean"]] + TBATSd[["mean"]])/5
#Combinationd2 <- (ETSd[["mean"]] + ETSd2[["mean"]] + ARIMAd[["mean"]] + ARIMAd2[["mean"]] +STLd[["mean"]] + MSTLd[["mean"]] + NNARd[["mean"]] + TBATSd[["mean"]])/8

autoplot(mytsd) +
  autolayer(ETSd, series="ETS", PI=FALSE) +
  autolayer(ARIMAd, series="ARIMA", PI=FALSE) +
  autolayer(STLd, series="STL", PI=FALSE) +
  autolayer(NNARd, series="NNAR", PI=FALSE) +
  autolayer(TBATSd, series="TBATS", PI=FALSE) +
  autolayer(Combinationd, series="Combination") +
  xlab("Year") + ylab("$") +
  ggtitle("Apple Daily Stock Price")
```

```{r}
#compare the RSME for all models calculated:
c(ETS = accuracy(ETSd, mytsd)["Test set","RMSE"],
  #ETS2 = accuracy(ETSd2, mytsd)["Test set","RMSE"],
  ARIMA = accuracy(ARIMAd, mytsd)["Test set","RMSE"],
  #ARIMA2 = accuracy(ARIMAd2, mytsd)["Test set","RMSE"],
  `STL-ETS` = accuracy(STLd, mytsd)["Test set","RMSE"],
  `MSTL-ETS` = accuracy(MSTLd, mytsd)["Test set","RMSE"],
  NNAR = accuracy(NNARd, mytsd)["Test set","RMSE"],
  TBATS = accuracy(TBATSd, mytsd)["Test set","RMSE"],
  Combination =accuracy(Combinationd, mytsd)["Test set","RMSE"]#,
  #Combination2 =accuracy(Combinationd2, mytsd)["Test set","RMSE"]
  )
```
NNAR does very well on its own, It beats the rest of them and the Conbination. 2nd Place goes to ARIMA.Since i am unaware of how to check residuals of a NNAR, we will do so for ARIMA
```{r}
mytsd.train %>% auto.arima(lambda = 0,biasadj = TRUE) %>% checkresiduals()
```

```{r}
#next weekly data
ETSw <- mytsw.train %>% ets() %>% forecast(h=hw)
ARIMAw <- mytsw.train %>% auto.arima(lambda = 0,biasadj = TRUE) %>% forecast(h=hw)
STLw <- mytsw.train  %>%  stlf(lambda = 0, h=hw, biasadj = TRUE)
#MSTLw <- mytsw.train  %>%  mstl(lambda = 0, biasadj = TRUE) %>% forecast(h=hw)
NNARw <- mytsw.train %>%  nnetar(lambda = 0) %>% forecast(h=hw)
TBATSw <- mytsw.train %>% tbats(biasadj = TRUE) %>% forecast(h=hw)
Combinationw <- (ETSw[["mean"]] + ARIMAw[["mean"]] +STLw[["mean"]] + NNARw[["mean"]] + TBATSw[["mean"]])/5

autoplot(mytsw) +
  autolayer(ETSw, series="ETS", PI=FALSE) +
  autolayer(ARIMAw, series="ARIMA", PI=FALSE) +
  autolayer(STLw, series="STL", PI=FALSE) +
  autolayer(NNARw, series="NNAR", PI=FALSE) +
  autolayer(TBATSw, series="TBATS", PI=FALSE) +
  autolayer(Combinationw, series="Combination") +
  xlab("Year") + ylab("$") +
  ggtitle("Apple Weekly Stock Price")
```

```{r}
#compare the RSME for all models calculated:
c(ETS = accuracy(ETSw, mytsw)["Test set","RMSE"],
  ARIMA = accuracy(ARIMAw, mytsw)["Test set","RMSE"],
  `STL-ETS` = accuracy(STLw, mytsw)["Test set","RMSE"],
  #`MSTL-ETS` = accuracy(MSTLw, mytsw)["Test set","RMSE"],
  NNAR = accuracy(NNARw, mytsw)["Test set","RMSE"],
  TBATS = accuracy(TBATSw, mytsw)["Test set","RMSE"],
  Combination =accuracy(Combinationw, mytsw)["Test set","RMSE"]
  )
```
Here again ARIMA does well yet this time its better than the Combination, and compared to the daily values the RSME has slighly increase for ARIMA.

```{r}
mytsw.train %>% auto.arima(lambda = 0,biasadj = TRUE) %>% checkresiduals()
```

```{r}
#lastly monthly data
ETSm <- mytsm.train %>% ets() %>% forecast(h=hm)
ARIMAm <- mytsm.train %>% auto.arima(lambda = 0,biasadj = TRUE) %>% forecast(h=hm)
STLm <- mytsm.train  %>%  stlf(lambda = 0, h=hm, biasadj = TRUE)
NNARm <- mytsm.train %>%  nnetar(lambda = 0) %>% forecast(h=hm)
TBATSm <- mytsm.train %>% tbats(biasadj = TRUE) %>% forecast(h=hm)
Combinationm <- (ETSm[["mean"]] + ARIMAm[["mean"]] +STLm[["mean"]] + NNARm[["mean"]] + TBATSm[["mean"]])/5

autoplot(mytsm) +
  autolayer(ETSm, series="ETS", PI=FALSE) +
  autolayer(ARIMAm, series="ARIMA", PI=FALSE) +
  autolayer(STLm, series="STL", PI=FALSE) +
  autolayer(NNARm, series="NNAR", PI=FALSE) +
  autolayer(TBATSm, series="TBATS", PI=FALSE) +
  autolayer(Combinationm, series="Combination") +
  xlab("Year") + ylab("$") +
  ggtitle("Apple Monthly Stock Price")
```

```{r}
#compare the RSME for all models calculated:
c(ETS = accuracy(ETSm, mytsm)["Test set","RMSE"],
  ARIMA = accuracy(ARIMAm, mytsm)["Test set","RMSE"],
  `STL-ETS` = accuracy(STLm, mytsm)["Test set","RMSE"],
  NNAR = accuracy(NNARm, mytsm)["Test set","RMSE"],
  TBATS = accuracy(TBATSm, mytsm)["Test set","RMSE"],
  Combination =accuracy(Combinationm, mytsm)["Test set","RMSE"]
  )
```
Here we see that ARIMA again is the leader even above the Combination again and that it has again increased from the Weekly and Daily values

```{r}
mytsm.train %>% auto.arima(lambda = 0,biasadj = TRUE) %>% checkresiduals()
```

```{r}
#mytsd.train %>% nsdiffs()
#mytsd.train %>% log() %>% ndiffs()
#mytsd.train %>% diff() %>% Box.test(type="Ljung-Box")
#library(urca)
#mytsd.train %>% ur.kpss() %>% summary()
#mytsd.train %>% diff() %>%  ur.kpss() %>% summary()

cbind("Original Data" = mytsd.train,
      "BoxCox" = BoxCox(mytsd.train, lambda = 0),
      "Seasonally\n differenced logs" =
        diff(log(mytsd.train),5),
      "Single\n differenced logs" =
        diff(BoxCox(mytsd.train, lambda = 0),1),
      "Double\n difference  logs" = 
        diff(diff(log(mytsd.train),5),1)) %>%
  autoplot(facets=TRUE) +
    xlab("Year") + ylab("") +
  ggtitle("Apple Daily Stock Price")
```

```{r}
#mytsw.train %>% nsdiffs()
#mytsw.train %>% log() %>% ndiffs()
#mytsw.train %>% diff() %>% Box.test(type="Ljung-Box")
#library(urca)
#mytsw.train %>% ur.kpss() %>% summary()
#mytsw.train %>% diff() %>%  ur.kpss() %>% summary()

cbind("Original Data" = mytsw.train,
      "BoxCox" = BoxCox(mytsw.train, lambda = 0),
      "Seasonally\n differenced logs" =
        diff(log(mytsw.train),52),
      "Single\n differenced logs" =
        diff(BoxCox(mytsw.train, lambda = 0),1),
      "Double\n difference  logs" = 
        diff(diff(log(mytsw.train),52),1)) %>%
  autoplot(facets=TRUE) +
    xlab("Year") + ylab("") +
  ggtitle("Apple Weekly Stock Price")
```

```{r}
#mytsm.train %>% nsdiffs()
#mytsm.train %>% log() %>% ndiffs()
#mytsm.train %>% diff() %>% Box.test(type="Ljung-Box")
#library(urca)
#mytsm.train %>% ur.kpss() %>% summary()
#mytsm.train %>% diff() %>%  ur.kpss() %>% summary()

cbind("Original Data" = mytsm.train,
      "BoxCox" = BoxCox(mytsm.train, lambda = 0),
      "Seasonally\n differenced logs" =
        diff(log(mytsm.train),12),
      "Single\n differenced logs" =
        diff(BoxCox(mytsm.train, lambda = 0),1),
      "Double\n difference  logs" = 
        diff(diff(log(mytsm.train),12),1)) %>%
  autoplot(facets=TRUE) +
    xlab("Year") + ylab("") +
  ggtitle("Apple Monthly Stock Price")
```

```{r}
#(2,1,2)(2,0,0) Daily
A1 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(2,0,0), lambda = 0)
A2 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(2,1,0), lambda = 0)
A3 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(2,1,1), lambda = 0)
A4 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(2,0,1), lambda = 0)
A5 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(2,0,2), lambda = 0)
A6 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(2,1,2), lambda = 0)
A7 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(1,1,1), lambda = 0)
A8 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(0,1,1), lambda = 0)
A9 <- mytsd.train %>% Arima(order = c(3,1,2), seasonal = c(2,0,0), lambda = 0)
A10 <- mytsd.train %>% Arima(order = c(2,1,3), seasonal = c(2,0,0), lambda = 0)
A11 <- mytsd.train %>% Arima(order = c(3,1,3), seasonal = c(2,0,0), lambda = 0)
A12 <- mytsd.train %>% Arima(order = c(4,1,2), seasonal = c(2,0,0), lambda = 0)
A13 <- mytsd.train %>% Arima(order = c(2,1,0), seasonal = c(2,0,0), lambda = 0)
A14 <- mytsd.train %>% Arima(order = c(2,1,0), seasonal = c(2,1,2), lambda = 0)


c(A1 = AIC(A1),
  A2 = AIC(A2),
  A3 = AIC(A3),
  A4 = AIC(A4),
  A5 = AIC(A5),
  A6 = AIC(A6),
  A7 = AIC(A7),
  A8 = AIC(A8),
  A9 = AIC(A9),
  A10 = AIC(A10),
  A11 = AIC(A11),
  A12 = AIC(A12),
  A13 = AIC(A13),
  A14 = AIC(A14)
)
```

```{r}
#(2,1,2)(2,0,0) Daily
A1 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hd)
A2 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(2,1,0), lambda = 0) %>% forecast(h=hd)
A3 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(2,1,1), lambda = 0) %>% forecast(h=hd)
A4 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(2,0,1), lambda = 0) %>% forecast(h=hd)
A5 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(2,0,2), lambda = 0) %>% forecast(h=hd)
A6 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(2,1,2), lambda = 0) %>% forecast(h=hd)
A7 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(1,1,1), lambda = 0) %>% forecast(h=hd)
A8 <- mytsd.train %>% Arima(order = c(2,1,2), seasonal = c(0,1,1), lambda = 0) %>% forecast(h=hd)
A9 <- mytsd.train %>% Arima(order = c(3,1,2), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hd)
A10 <- mytsd.train %>% Arima(order = c(2,1,3), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hd)
A11 <- mytsd.train %>% Arima(order = c(3,1,3), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hd)
A12 <- mytsd.train %>% Arima(order = c(4,1,2), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hd)
A13 <- mytsd.train %>% Arima(order = c(2,1,0), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hd)
A14 <- mytsd.train %>% Arima(order = c(2,1,0), seasonal = c(2,1,2), lambda = 0) %>% forecast(h=hd)

#compare the RSME for all models calculated:
c(A1 = accuracy(A1, mytsd)["Test set","RMSE"],
  A2 = accuracy(A2, mytsd)["Test set","RMSE"],
  A3 = accuracy(A3, mytsd)["Test set","RMSE"],
  A4 = accuracy(A4, mytsd)["Test set","RMSE"],
  A5 = accuracy(A5, mytsd)["Test set","RMSE"],
  A6 = accuracy(A6, mytsd)["Test set","RMSE"],
  A7 = accuracy(A7, mytsd)["Test set","RMSE"],
  A8 = accuracy(A8, mytsd)["Test set","RMSE"],
  A9 = accuracy(A9, mytsd)["Test set","RMSE"],
  A10 = accuracy(A10, mytsd)["Test set","RMSE"],
  A11 = accuracy(A11, mytsd)["Test set","RMSE"],
  A12 = accuracy(A12, mytsd)["Test set","RMSE"],
  A13 = accuracy(A13, mytsd)["Test set","RMSE"],
  A14 =accuracy(A14, mytsd)["Test set","RMSE"]
  )
```

```{r}
#(3,1,1)(2,0,1) Weekly
A1  <- mytsw.train %>% auto.arima(lambda = 0,biasadj = TRUE)
#A2 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(2,1,0), lambda = 0)
#A3 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(2,1,1), lambda = 0)
#A4 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(2,0,0), lambda = 0)
#A5 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(2,0,2), lambda = 0)
#A6 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(2,1,2), lambda = 0)
A7 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(1,1,1), lambda = 0)
A8 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(0,1,1), lambda = 0)
#A9 <- mytsw.train %>% Arima(order = c(3,1,2), seasonal = c(2,0,0), lambda = 0)
#A10 <- mytsw.train %>% Arima(order = c(2,1,1), seasonal = c(2,0,0), lambda = 0)
A11 <- mytsw.train %>% Arima(order = c(3,1,3), seasonal = c(2,0,0), lambda = 0)
#A12 <- mytsw.train %>% Arima(order = c(4,1,2), seasonal = c(2,0,0), lambda = 0)
#A13 <- mytsw.train %>% Arima(order = c(3,1,0), seasonal = c(2,0,0), lambda = 0)
#A14 <- mytsw.train %>% Arima(order = c(3,1,0), seasonal = c(2,1,2), lambda = 0)


c(A1 = AIC(A1),
  #A2 = AIC(A2),
  #A3 = AIC(A3),
  #A4 = AIC(A4),
  #A5 = AIC(A5),
  #A6 = AIC(A6),
  A7 = AIC(A7),
  A8 = AIC(A8),
  #A9 = AIC(A9),
  #A10 = AIC(A10),
  A11 = AIC(A11)#,
  #A12 = AIC(A12),
  #A13 = AIC(A13),
  #A14 = AIC(A14)
)
```

```{r}
#(3,1,1)(2,0,1) Weekly %>% forecast(h=hw)
A1  <- ARIMAw
#A2 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(2,1,0), lambda = 0) %>% forecast(h=hw)
#A3 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(2,1,1), lambda = 0) %>% forecast(h=hw)
#A4 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hw)
#A5 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(2,0,2), lambda = 0) %>% forecast(h=hw)
#A6 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(2,1,2), lambda = 0) %>% forecast(h=hw)
A7 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(1,1,1), lambda = 0) %>% forecast(h=hw)
A8 <- mytsw.train %>% Arima(order = c(3,1,1), seasonal = c(0,1,1), lambda = 0) %>% forecast(h=hw)
#A9 <- mytsw.train %>% Arima(order = c(3,1,2), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hw)
#A10 <- mytsw.train %>% Arima(order = c(2,1,1), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hw)
A11 <- mytsw.train %>% Arima(order = c(3,1,3), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hw)
#A12 <- mytsw.train %>% Arima(order = c(4,1,2), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hw)
#A13 <- mytsw.train %>% Arima(order = c(3,1,0), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hw)
#A14 <- mytsw.train %>% Arima(order = c(3,1,0), seasonal = c(2,1,2), lambda = 0) %>% forecast(h=hw)

#compare the RSME for all models calculated:
c(A1 = accuracy(A1, mytsw)["Test set","RMSE"],
  #A2 = accuracy(A2, mytsd)["Test set","RMSE"],
  #A3 = accuracy(A3, mytsd)["Test set","RMSE"],
  #A4 = accuracy(A4, mytsd)["Test set","RMSE"],
  #A5 = accuracy(A5, mytsd)["Test set","RMSE"],
  #A6 = accuracy(A6, mytsd)["Test set","RMSE"],
  A7 = accuracy(A7, mytsw)["Test set","RMSE"],
  A8 = accuracy(A8, mytsw)["Test set","RMSE"],
  #A9 = accuracy(A9, mytsd)["Test set","RMSE"],
  #A10 = accuracy(A10, mytsd)["Test set","RMSE"],
  A11 = accuracy(A11, mytsw)["Test set","RMSE"]#,
  #A12 = accuracy(A12, mytsd)["Test set","RMSE"],
  #A13 = accuracy(A13, mytsd)["Test set","RMSE"],
  #A14 =accuracy(A14, mytsd)["Test set","RMSE"]
  )
```

```{r}
#(0,1,0)(2,0,0) Monthly
A1 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(2,0,0), lambda = 0) 
A2 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(2,1,0), lambda = 0) 
A3 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(2,1,1), lambda = 0) 
A4 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(2,0,1), lambda = 0) 
A5 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(2,0,2), lambda = 0) 
A6 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(2,1,2), lambda = 0) 
A7 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(1,1,1), lambda = 0) 
A8 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(0,1,1), lambda = 0) 
A9 <- mytsm.train %>% Arima(order = c(1,1,1), seasonal = c(2,0,0), lambda = 0) 
A10 <- mytsm.train %>% Arima(order = c(0,1,1), seasonal = c(2,0,0), lambda = 0) 
A11 <- mytsm.train %>% Arima(order = c(1,1,0), seasonal = c(2,0,0), lambda = 0)
A12 <- mytsm.train %>% Arima(order = c(0,1,2), seasonal = c(2,0,0), lambda = 0) 
A13 <- mytsm.train %>% Arima(order = c(1,1,2), seasonal = c(2,0,0), lambda = 0) 
A14 <- mytsm.train %>% Arima(order = c(0,2,0), seasonal = c(2,0,0), lambda = 0) 

c(A1 = AIC(A1),
  A2 = AIC(A2),
  A3 = AIC(A3),
  A4 = AIC(A4),
  A5 = AIC(A5),
  A6 = AIC(A6),
  A7 = AIC(A7),
  A8 = AIC(A8),
  A9 = AIC(A9),
  A10 = AIC(A10),
  A11 = AIC(A11),
  A12 = AIC(A12),
  A13 = AIC(A13),
  A14 = AIC(A14)
)
```

```{r}
#(0,1,0)(2,0,0) Monthly
A1 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hm)
A2 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(2,1,0), lambda = 0) %>% forecast(h=hm)
A3 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(2,1,1), lambda = 0) %>% forecast(h=hm)
A4 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(2,0,1), lambda = 0) %>% forecast(h=hm)
A5 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(2,0,2), lambda = 0) %>% forecast(h=hm)
A6 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(2,1,2), lambda = 0) %>% forecast(h=hm)
A7 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(1,1,1), lambda = 0) %>% forecast(h=hm)
A8 <- mytsm.train %>% Arima(order = c(0,1,0), seasonal = c(0,1,1), lambda = 0) %>% forecast(h=hm)
A9 <- mytsm.train %>% Arima(order = c(1,1,1), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hm)
A10 <- mytsm.train %>% Arima(order = c(0,1,1), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hm)
A11 <- mytsm.train %>% Arima(order = c(1,1,0), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hm)
A12 <- mytsm.train %>% Arima(order = c(0,1,2), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hm)
A13 <- mytsm.train %>% Arima(order = c(1,1,0), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hm)
A14 <- mytsm.train %>% Arima(order = c(0,2,0), seasonal = c(2,0,0), lambda = 0) %>% forecast(h=hm)

#compare the RSME for all models calculated:
c(A1 = accuracy(A1, mytsm)["Test set","RMSE"],
  A2 = accuracy(A2, mytsm)["Test set","RMSE"],
  A3 = accuracy(A3, mytsm)["Test set","RMSE"],
  A4 = accuracy(A4, mytsm)["Test set","RMSE"],
  A5 = accuracy(A5, mytsm)["Test set","RMSE"],
  A6 = accuracy(A6, mytsm)["Test set","RMSE"],
  A7 = accuracy(A7, mytsm)["Test set","RMSE"],
  A8 = accuracy(A8, mytsm)["Test set","RMSE"],
  A9 = accuracy(A9, mytsm)["Test set","RMSE"],
  A10 = accuracy(A10, mytsm)["Test set","RMSE"],
  A11 = accuracy(A11, mytsm)["Test set","RMSE"],
  A12 = accuracy(A12, mytsm)["Test set","RMSE"],
  A13 = accuracy(A13, mytsm)["Test set","RMSE"],
  A14 =accuracy(A14, mytsm)["Test set","RMSE"]
  )
```


