---
title: "Unit 2"
output: html_notebook
---

2.4 problem loading data from CSV file, that was previously imported

2.4 Suppose that a hospital tested the age and body fat data for 18 randomly selected adults
with the following results:
age 23 23 27 27 39 41 47 49 50
%fat 9.5 26.5 7.8 17.8 31.4 25.9 27.4 27.2 31.2
age 52 54 54 56 57 58 58 60 61
%fat 34.6 42.5 28.8 33.4 30.2 34.1 32.9 41.2 35.7
```{r}
age <-c (`2.4problem`$age)
age
xfat <- c(`2.4problem`$X.fat)
xfat
```
(a) Calculate the mean, median, and standard deviation of age and %fat.
```{r}
# Create the function.
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

mean(age)
median(age)
getmode(age)

mean(xfat)
median(xfat)
getmode(xfat)
```
(b) Draw the boxplots for age and %fat.
```{r}
boxplot(age,xfat)
```
2.6 Given two objects represented by the tuples (22, 1, 42, 10) and (20, 0, 36, 8):
```{r}
t1 <-c(22,1,42,10)
t1m <-matrix(c(22,1,42,10),nrow=1)
t1
t1m
t2 <-c(20,0,36,8)
t2m <-matrix(c(20,0,36,8),nrow=1)
t2
t2m
```
(a) Compute the Euclidean distance between the two objects.
manually first:
```{r}
a0 <- t1m-t2m
a0
a1 <- a0^2
a1
a2 <- sum(a1[1],a1[2],a1[3],a1[4])
a2
a3 <- sqrt(a2)
a3
```
using a function:
```{r}
dist(rbind(t1,t2))
```

wiring my own function:
```{r}
euc.dist <- function(x1, x2) sqrt(sum((x1 - x2) ^ 2))
euc.dist(t1,t2)
euc.dist(t1m,t2m)
```
(b) Compute the Manhattan distance between the two objects.
manually:
```{r}
a0 <- abs(t1m-t2m)
a0
a1 <- sum(a0[1],a0[2],a0[3],a0[4])
a1
```
own function:
```{r}
manhattan.dist <- function(x1,x2) sum(abs(x1-x2))
manhattan.dist(t1,t2)
```
function:
```{r}
dist(rbind(t1,t2), method = "manhattan")
```
(c) Compute the Minkowski distance between the two objects, using q=3.
own function:
```{r}
minkowski.dist <- function(x1, x2,q) (sum(abs(x1 - x2) ^ q))^(1/q)
minkowski.dist(t1m,t2m,3)
```
(d) Compute the supremum distance between the two objects.
function:
```{r}
dist(rbind(t1,t2),method = "maximum")
```

Chaz Clark
2/9/19
