---
title: "Feature Engineering using R"
output: html_notebook
name: Chaz Clark MDS 556 Week 3
source: Feature Engineering using R
https://techcommunity.microsoft.com/t5/SQL-Server/Feature-Engineering-using-R/ba-p/385286
https://blogs.msdn.microsoft.com/microsoftrservertigerteam/2017/03/23/feature-engineering-using-r/
---

This will load all the available datsets built into R:
```{r}
#data()
head(iris)
head(mtcars)
```

Plot the Pearson Correlation:
```{r}
plot(mtcars$wt, mtcars$mpg, col="red", xlab = "Weight of the cars", ylab = "Miles per gallon", pch=16, main = "Pearson Correlation")
```

Requires Microsoft R Open:
```{r}
#form <- ~ wt + mpg + cyl
#rxCor(form, data = mtcars)
```

Look at the correlation of the two fields:
```{r}
cor.test(x = mtcars$wt, y = mtcars$mpg)
```

Plot the pairs to view correlation:
```{r}
pairs(~mpg + disp + wt + cyl, data=mtcars, main="Scatterplot Matrix")
```

Look at the IG Values of the IRIS data:
```{r}
library(FSelectorRcpp)
ig_values <- information_gain(Species ~ ., iris)
cut_attrs(attrs = x)
ig_values
```

```{r}
# filter the top K features and use it to create the model formula
top_k_features <- to_formula(cut_attrs(attrs = x), "Species")
f <- cut_attrs(attrs = x, k = 2)
f
```

Needs Microsoft R Server:
```{r}
#logistic_Mod <- rxLogisticRegression(Species ~ Petal.Width + Petal.Length + Sepal.Width + Sepal.Length, data = iris, l1Weight = 0, 
#type = "multi", mlTransforms = list(selectFeatures(Species ~ Petal.Width + Petal.Length + Sepal.Width + Sepal.Length, mode = mutualInformation(numFeaturesToKeep = 2))))
```

Wrapper Method:
```{r}
library(caret)
rfe_controller <- rfeControl(functions=lmFuncs, method="repeatedcv", repeats = 5, verbose = FALSE)
size <- c(1:10) 
lm_Profiler <- rfe(mtcars[,2:11], mtcars[,1], rfeControl = rfe_controller)
lm_Profiler
```

```{r}
plot(lm_Profiler)
```

Feature Extraction or Construction:
```{r}
range(CO2$conc)
scaled_conc <- scale(CO2$conc)
range(scaled_conc)
```

PCA:
```{r}
mtcars_feat <- mtcars[,2:11]
mtcars_mpg <- mtcars[,1]
mtcars_pca <- prcomp(x = mtcars_feat, scale. = T)
# gives the standard deviation of all the PCAs.
mtcars_pca$sdev
```

```{r}
# List of PCAs generated for the mtcars dataset
mtcars_pca
```

Needs Microsoft R Server:
```{r}
#Use Biplot to visualize the first two principal components:
#biplot(mtcars.pca_model, scale = TRUE, pc.biplot = TRUE)
```

LDA: 
```{r}
#head(iris)
#train <- sample(1:150, 75)
#lda_model <- lda(Species ~., data = iris, subset=train)
```

```{r}
# shows you the mean, used for LDA
 #lda_model$means
```

```{r}
#Predictions on the test data
#lda_pred <- predict(object = lda_model, newdata = iris[-train, ])
#lda_pred$class
#iris_test <- iris[-train, ]
#ggplot() + geom_point(aes(lda_pred$x[,1], lda_pred$x[,2], colour = iris_test$Species, shape = iris_test$Species), 
#size = 2.5) + ggtitle("Linear Discriminant Analysis") + xlab("LDA1") + ylab("LDA2") + labs(fill = "Species")
```

package ‘MicrosoftML’ is not available (for R version 3.6.1)
```{r}
#library(MicrosoftML)
#rf_model <- rxDForest(formula = form, data = mtcars, importance = TRUE )
#rf_model$importance
```
```{r}
#rf_model <- randomForest(mpg ~ ., data = mtcars, importance = TRUE)
#rf_model$importance
```

Random Forest:
```{r}
# NOT RUN {
library(randomForest)
set.seed(4543)
data(mtcars)
mtcars.rf <- randomForest(mpg ~ ., data=mtcars, ntree=1000,
                          keep.forest=FALSE, importance=TRUE)
importance(mtcars.rf)
importance(mtcars.rf, type=1)
# }
```

