---
title: "Data Assignment: Random Numbers and Correlation"
output: html_notebook
name: Chaz Clark MDS 556 Fall 2019
---
---
assignment:
The file Some Random Numbers to Play With contains 20,000 random numbers generated by a high quality method from Random.org. They have been arranged, maintaining full randomness, into 1,000 rows with 20 columns in each row.

These data can be viewed in two fundamentally different ways. One approach is to consider them as twenty columns or features with 1000 data points for each. The other is to view them as 1,000 rows or features with 20 data points each.

Explore the behavior of correlations using each of the two views. Describe any patterns or surprises you observe in these correlations, especially given that these data are highly random. How do your observations fir with Granville's first article?
---

Import the data 1000 rows x 20 columns Dataset

```{r}
#Random_data_v2 <- read_excel("Documents/Documents - Mac Pro/Elmhurst/MDS 556 Analytical #Methods/Week1/Random data v2.xlsx", 
#+     sheet = "Sheet1", col_names = FALSE)
```

load data into a matrix

```{r}
res1<-rcorr(as.matrix(y[,1:20]))
```

load a flatten function:

```{r}
# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
#http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
```

Display the flattened results:

```{r}
flattenCorrMatrix(res1$r, res1$P)
```

Import the data 20 rows x 1000 columns Dataset (copy paste transpose in excel)

```{r}
#Random_data_v2_1 <- read_excel("Documents/Documents - Mac Pro/Elmhurst/MDS 556 Analytical #Methods/Week1/Random data v2.1.xlsx", 
#+     sheet = "Sheet3", col_names = FALSE)
```

load data into a matrix2

```{r}
res2<-rcorr(as.matrix(y[,1:1000]))
```

Display the flattened results2:

```{r}
flattenCorrMatrix(res2$r, res2$P)
```

Q.How do your observations fir with Granville's first article?

A.
I am not seeing the correlation within the random dataset in its natural form. I am not adding any feature engineering but in both datasets (not looking over all 100 pages), having sampled multiple pages no correlation is close to 1 and no p values are near a significant threshold. I am just not seeing what Granville is talking about with hidden patterns in the data. Perhaps the data is not Big enough data to be big data at this point.












Notebook - how to use notes:

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.